---
title: \vspace{-1cm} Part7_Raw_Analysis_DESEQ_Fig7-13 \vspace{-0.20cm}
author: "Julien G. Roth"
date: "3/15/2021"
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
output: pdf_document
---

# Performing DESEQ Analysis #

DESEQ is used to perform statistical comparisons of various groups.

Input: **data_x_m_f_neuralECM_f_counts**, **data_x_m_f_neuralECM_metadata**

Output: 

Key: x (Trevino study), m (mean expression), f (timepoint filtered), neuralECM (neural ECM gene filtered), f (low expression filtered)

**Set-up DESEQ design parameters**

```{r, tidy=TRUE}

design_celltype <- as.formula(~Cell_Type)
design_maturity <- as.formula(~Maturity)
design_patterning <- as.formula(~Patterning)
```

**DESEQ comparing cell types**

```{r, tidy=TRUE}

ddsobj_celltype <- DESeqDataSetFromMatrix(countData = round(data_x_m_f_neuralECM_f_counts),
                              colData = data_x_m_f_neuralECM_metadata,
                              design = design_celltype)

dds_celltype <- DESeq(ddsobj_celltype)

vsd_celltype <- varianceStabilizingTransformation(dds_celltype, blind=FALSE)
head(assay(vsd_celltype), 3)
```

**Results neuronal vs glial comparisons**

```{r, tidy=TRUE}

res_celltype_ng <- results(dds_celltype, c("Cell_Type", "neuronal", "glial"))

summary(res_celltype_ng)
```

**Plot normalized count for gene with the lowest p-value when comparing neuronal to glial (whole is shown)**

```{r}

plotCounts(dds_celltype, gene=which.min(res_celltype_ng$padj), intgroup="Cell_Type")
```

**Shrink log2 fold changes and create volcano plot**

```{r, tidy=TRUE}

res_celltype_ng_shr <- lfcShrink(dds_celltype, contrast = c('Cell_Type','neuronal','glial'), res=res_celltype_ng, type = 'normal')

EnhancedVolcano(res_celltype_ng_shr,
    lab = rownames(res_celltype_ng_shr),
    x = 'log2FoldChange',
    y = 'pvalue',
    pCutoff = 10e-5,
    FCcutoff = 1.5,
    colAlpha = 0.5,
    pointSize = 3.0,
    labSize = 3.0,
    labvjust = 0.01,
    labhjust = -1,
    title = "Figure 7.) Volcano Plot of Neural ECM Gene Expression",
    subtitle = "Neuronal vs Glial Cells",
    titleLabSize = 14,
    subtitleLabSize = 12,
    caption = bquote(~Log[2]~ "fold change cutoff, 1.5; p-value cutoff, 10e-5"),
    captionLabSize = 10,
    axisLabSize = 12,
    legendPosition = "right",
    legendLabSize = 10,
    drawConnectors = TRUE,
    widthConnectors = 0.5,
    gridlines.major = FALSE,
    gridlines.minor = FALSE,
    cutoffLineType = 'twodash',
    cutoffLineWidth = 0.5,)

ggsave('figure_7f_raw_neuralECMgenes_DESEQ_ng_volcano.pdf', 
       device = "pdf", 
       path = str_c(directory, folder_fig_r), 
       width=12, 
       height=6, 
       units = "in", 
       scale = 1)
```

**Create heatmaps of comparison**

```{r, tidy=TRUE}

color_palette <- colorRampPalette(c("#ffffff", "#deebf7","#9ecae1","#4292c6","#08519c"))(n = 50)

ann_colors = list(Patterning = c(hCS = "#252525", hSS = "#d9d9d9"), 
                  Maturity = c(early = "#d9d9d9", intermediate = "#969696", late = "#252525"),
                  Cell_Type = c(whole = "#969696", neuronal = "#252525", glial = "#bdbdbd"))

ng_topVarGenes <- head(order(rowVars(assay(dds_celltype)), decreasing = TRUE), 20)
ng_mat  <- assay(dds_celltype)[ng_topVarGenes, ]
ng_mat  <- ng_mat - rowMeans(ng_mat)
ng_anno <- as.data.frame(colData(dds_celltype)[, c("Cell_Type","Maturity", "Patterning")])

pheatmap(ng_mat,
         annotation_col = ng_anno,
         cluster_rows = FALSE, 
         cluster_cols = TRUE, 
         scale = 'none',
         cutree_cols = 4,
         treeheight_col = 15,
         fontsize_row = 12,
         fontsize_col = 16,
         color = color_palette,
         annotation_colors = ann_colors,
         border_color = "NA",
         show_colnames = TRUE,
         angle_col = 90,
         main = "Figure 8.) Heatmap of Neural ECM Gene Expression", 
         width = 12, 
         height = 12,
         filename = str_c(directory, folder_fig_r, "/figure_8_raw_neuralECMgenes_DESEQ_ng_heatmap.pdf"))

vsd_ng_topVarGenes <- head(order(rowVars(assay(vsd_celltype)), decreasing = TRUE), 20)
vsd_ng_mat  <- assay(vsd_celltype)[vsd_ng_topVarGenes, ]
vsd_ng_mat  <- vsd_ng_mat - rowMeans(vsd_ng_mat)
vsd_ng_anno <- as.data.frame(colData(vsd_celltype)[, c("Cell_Type","Maturity", "Patterning")])

pheatmap(vsd_ng_mat,
         annotation_col = vsd_ng_anno,
         cluster_rows = FALSE, 
         cluster_cols = TRUE, 
         scale = 'none',
         cutree_cols = 4,
         treeheight_col = 15,
         fontsize_row = 12,
         fontsize_col = 16,
         color = color_palette,
         annotation_colors = ann_colors,
         border_color = "NA",
         show_colnames = TRUE,
         angle_col = 90,
         main = "Figure 9.) Heatmap of VSD of Neural ECM Gene Expression", 
         width = 12, 
         height = 12,
         filename = str_c(directory, folder_fig_r, "/figure_9_raw_neuralECMgenes_DESEQ_VSD_ng_heatmap.pdf"))
```


\pagebreak


**DESEQ comparing maturity**

```{r, tidy=TRUE}

ddsobj_maturity <- DESeqDataSetFromMatrix(countData = round(data_x_m_f_neuralECM_f_counts),
                              colData = data_x_m_f_neuralECM_metadata,
                              design = design_maturity)

dds_maturity <- DESeq(ddsobj_maturity)

vsd_maturity <- varianceStabilizingTransformation(dds_maturity, blind=FALSE)
head(assay(vsd_maturity), 3)
```

**Results neuronal vs glial comparisons**

```{r, tidy=TRUE}

res_mat_il <- results(dds_maturity, c("Maturity", "intermediate", "late"))

summary(res_mat_il)
```

**Plot normalized count for gene with the lowest p-value when comparing neuronal to glial (whole is shown)**

```{r}

plotCounts(dds_maturity, gene=which.min(res_mat_il$padj), intgroup="Maturity")
```

**Shrink log2 fold changes and create volcano plot**

```{r, tidy=TRUE}

res_mat_il_shr <- lfcShrink(dds_maturity, contrast = c('Maturity','intermediate','late'), res=res_mat_il, type = 'normal')

EnhancedVolcano(res_mat_il_shr,
    lab = rownames(res_mat_il_shr),
    x = 'log2FoldChange',
    y = 'pvalue',
    pCutoff = 10e-5,
    FCcutoff = 1.5,
    colAlpha = 0.5,
    pointSize = 3.0,
    labSize = 3.0,
    labvjust = 0.01,
    labhjust = -1,
    title = "Figure 10.) Volcano Plot of Neural ECM Gene Expression",
    subtitle = "Intermediate vs Late",
    titleLabSize = 14,
    subtitleLabSize = 12,
    caption = bquote(~Log[2]~ "fold change cutoff, 1.5; p-value cutoff, 10e-5"),
    captionLabSize = 10,
    axisLabSize = 12,
    legendPosition = "right",
    legendLabSize = 10,
    drawConnectors = TRUE,
    widthConnectors = 0.5,
    gridlines.major = FALSE,
    gridlines.minor = FALSE,
    cutoffLineType = 'twodash',
    cutoffLineWidth = 0.5,)

ggsave('figure_10_raw_neuralECMgenes_DESEQ_il_volcano.pdf', 
       device = "pdf", 
       path = str_c(directory, folder_fig_r), 
       width=12, 
       height=6, 
       units = "in", 
       scale = 1)
```

**Create heatmap of comparison**

```{r}
vsd_mat_topVarGenes <- head(order(rowVars(assay(vsd_maturity)), decreasing = TRUE), 20)
vsd_mat_mat  <- assay(vsd_maturity)[vsd_mat_topVarGenes, ]
vsd_mat_mat  <- vsd_mat_mat - rowMeans(vsd_mat_mat)
vsd_mat_anno <- as.data.frame(colData(vsd_maturity)[, c("Cell_Type","Maturity", "Patterning")])

pheatmap(vsd_mat_mat,
         annotation_col = vsd_mat_anno,
         cluster_rows = FALSE, 
         cluster_cols = TRUE, 
         scale = 'none',
         cutree_cols = 4,
         treeheight_col = 15,
         fontsize_row = 12,
         fontsize_col = 16,
         color = color_palette,
         annotation_colors = ann_colors,
         border_color = "NA",
         show_colnames = TRUE,
         angle_col = 90,
         main = "Figure 11.) Heatmap of VSD of Neural ECM Gene Expression", 
         width = 12, 
         height = 12,
         filename = str_c(directory, folder_fig_r, "/figure_11_raw_neuralECMgenes_DESEQ_VSD_mat_heatmap.pdf"))
```


\pagebreak


**DESEQ comparing patterning**

```{r, tidy=TRUE}

ddsobj_patterning <- DESeqDataSetFromMatrix(countData = round(data_x_m_f_neuralECM_f_counts),
                              colData = data_x_m_f_neuralECM_metadata,
                              design = design_patterning)

dds_patterning <- DESeq(ddsobj_patterning)

vsd_patterning <- varianceStabilizingTransformation(dds_patterning, blind=FALSE)
head(assay(vsd_patterning), 3)
```

**Results hCS vs hSS comparisons**

```{r, tidy=TRUE}

res_patterning <- results(dds_patterning, c("Patterning", "hCS", "hSS"))

summary(res_patterning)
```

**Plot normalized count for gene with the lowest p-value when comparing neuronal to glial (whole is shown)**

```{r}

plotCounts(dds_patterning, gene=which.min(res_patterning$padj), intgroup="Patterning")
```

**Shrink log2 fold changes and create volcano plot**

```{r, tidy=TRUE}

res_patterning_shr <- lfcShrink(dds_patterning, contrast = c('Patterning','hCS','hSS'), res=res_patterning, type = 'normal')

EnhancedVolcano(res_patterning_shr,
    lab = rownames(res_patterning_shr),
    x = 'log2FoldChange',
    y = 'pvalue',
    pCutoff = 10e-5,
    FCcutoff = 1.5,
    colAlpha = 0.5,
    pointSize = 3.0,
    labSize = 3.0,
    labvjust = 0.01,
    labhjust = -1,
    title = "Figure 12.) Volcano Plot of Neural ECM Gene Expression",
    subtitle = "hCS vs hSS",
    titleLabSize = 14,
    subtitleLabSize = 12,
    caption = bquote(~Log[2]~ "fold change cutoff, 1.5; p-value cutoff, 10e-5"),
    captionLabSize = 10,
    axisLabSize = 12,
    legendPosition = "right",
    legendLabSize = 10,
    drawConnectors = TRUE,
    widthConnectors = 0.5,
    gridlines.major = FALSE,
    gridlines.minor = FALSE,
    cutoffLineType = 'twodash',
    cutoffLineWidth = 0.5,)

ggsave('figure_12_raw_neuralECMgenes_DESEQ_pat_volcano.pdf', 
       device = "pdf", 
       path = str_c(directory, folder_fig_r), 
       width=12, 
       height=6, 
       units = "in", 
       scale = 1)
```

**Create heatmap of comparison**

```{r}
vsd_pat_topVarGenes <- head(order(rowVars(assay(vsd_patterning)), decreasing = TRUE), 20)
vsd_pat_mat  <- assay(vsd_patterning)[vsd_pat_topVarGenes, ]
vsd_pat_mat  <- vsd_pat_mat - rowMeans(vsd_pat_mat)
vsd_pat_anno <- as.data.frame(colData(vsd_patterning)[, c("Cell_Type","Maturity", "Patterning")])

pheatmap(vsd_pat_mat,
         annotation_col = vsd_pat_anno,
         cluster_rows = FALSE, 
         cluster_cols = TRUE, 
         scale = 'none',
         cutree_cols = 4,
         treeheight_col = 15,
         fontsize_row = 12,
         fontsize_col = 16,
         color = color_palette,
         annotation_colors = ann_colors,
         border_color = "NA",
         show_colnames = TRUE,
         angle_col = 90,
         main = "Figure 13.) Heatmap of VSD of Neural ECM Gene Expression", 
         width = 12, 
         height = 12,
         filename = str_c(directory, folder_fig_r, "/figure_13_raw_neuralECMgenes_DESEQ_VSD_pat_heatmap.pdf"))
```


\pagebreak


# Cleaning Data #

**Export the variables which are requried for downstream steps.**

```{r, tidy=TRUE}

write.csv(as.data.frame(res_celltype_ng), str_c(directory, variables, "data_x_m_f_neuralECM_f_DESEQ_ng.csv"))
write.csv(as.data.frame(res_mat_il), str_c(directory, variables, "data_x_m_f_neuralECM_f_DESEQ_il.csv"))
write.csv(as.data.frame(res_patterning), str_c(directory, variables, "data_x_m_f_neuralECM_f_DESEQ_pat.csv"))
```

**Remove less relevant or temporary variables to decrease clutter.**

```{r, tidy=TRUE}

rm(ann_colors, color_palette, design_celltype, design_maturity, design_patterning, ng_topVarGenes, ng_mat, ng_anno, res_celltype_ng_shr, res_mat_il_shr, res_patterning_shr, vsd_mat_anno, vsd_mat_mat, vsd_ng_anno, vsd_ng_mat, vsd_pat_anno, vsd_pat_mat)
```