---
title: \vspace{-1cm} Part3_A_Raw_DataProcessing_Aggregate_Time \vspace{-0.20cm}
author: "Julien G. Roth"
date: "3/14/2021"
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
output: pdf_document
---

# Filtering Data by Time # 

**The following script attempts to reduce some of the dimensionality of the dataset. This is accomplished by aggregating the data into one of the following categories: Early Whole, Intermediate Neuronal, Intermediate Glial, Late Neuronal, Late Glial. Unfortunately glial purification was only performed once in hSS patterning.**   

Some timepoints are excluded if both neuronal/glial cells were not sampled.  

The output is the dataframe: **'data_x_m_f'**. The input file, 'data_x_m' still has all timepoints for each condition.**  


```{r, tidy=TRUE}

data_x_m_filter_hCS_w <- data_x_m %>% 
  filter(Patterning == "hCS") %>% 
  filter(Cell_Type == "whole") %>% 
  mutate(Maturity = "early")

data_x_m_filter_hCS_n <- data_x_m %>% 
  filter(Patterning == "hCS") %>% 
  filter(Cell_Type == "neuronal") %>% 
  filter(Day == 140 | Day == 179 | Day == 206 | Day == 220 | Day == 588 | Day == 595) %>% 
  mutate(Maturity = case_when(Day == 140 ~ "intermediate",  Day == 179 ~ "intermediate", Day == 206 ~ "intermediate", Day == 220 ~ "intermediate", 
                              Day == 588 ~ "late", Day == 595 ~ "late"))

data_x_m_filter_hCS_g <- data_x_m %>% 
  filter(Patterning == "hCS") %>% 
  filter(Cell_Type == "glial") %>% 
  filter(Day == 140 | Day == 179 | Day == 206 | Day == 220 | Day == 588 | Day == 595) %>% 
  mutate(Maturity = case_when(Day == 140 ~ "intermediate",  Day == 179 ~ "intermediate", Day == 206 ~ "intermediate", Day == 220 ~ "intermediate", 
                              Day == 588 ~ "late", Day == 595 ~ "late"))

data_x_m_filter_hCS <- rbind(data_x_m_filter_hCS_w, data_x_m_filter_hCS_n, data_x_m_filter_hCS_g)

data_x_m_filter_hCS <- data_x_m_filter_hCS %>% 
  mutate(Cell_Category = paste(Maturity, Cell_Type, sep = "-"))

head(data_x_m_filter_hCS, 5)
```


\pagebreak


```{r, tidy=TRUE}

data_x_m_filter_hSS_w <- data_x_m %>% 
  filter(Patterning == "hSS") %>% 
  filter(Cell_Type == "whole") %>% 
  mutate(Maturity = "early")

data_x_m_filter_hSS_n <- data_x_m %>% 
  filter(Patterning == "hSS") %>% 
  filter(Cell_Type == "neuronal") %>% 
  filter(Day == 125 | Day == 588) %>% 
  mutate(Maturity = case_when(Day == 125 ~ "intermediate", Day == 588 ~ "late"))

data_x_m_filter_hSS_g <- data_x_m %>% 
  filter(Patterning == "hSS") %>% 
  filter(Cell_Type == "glial") %>% 
  filter(Day == 125) %>% 
  mutate(Maturity = "intermediate")

data_x_m_filter_hSS <- rbind(data_x_m_filter_hSS_w, data_x_m_filter_hSS_n, data_x_m_filter_hSS_g)

data_x_m_filter_hSS <- data_x_m_filter_hSS %>% 
  mutate(Cell_Category = paste(Maturity, Cell_Type, sep = "-"))

head(data_x_m_filter_hSS, 5)
```


\pagebreak


**Compiling the resultant time-filtered dataframes.**

```{r, tidy=TRUE}

data_x_m_filter <- rbind(data_x_m_filter_hCS, data_x_m_filter_hSS)

data_x_m_f <- data_x_m_filter

nrow(data_x_m_f)
```


\pagebreak


# Cleaning Data #

**Export the variables which are requried for downstream steps.**

```{r, tidy=TRUE}

write.table(data_x_m_f, str_c(directory, variables, "data_x_m_f.csv"), sep = ',', row.names = FALSE)
```

**Remove less relevant or temporary variables to decrease clutter.**

```{r, tidy=TRUE}

rm(data_x_m_filter, data_x_m_filter_hCS, data_x_m_filter_hCS_g, data_x_m_filter_hCS_n, data_x_m_filter_hCS_w, data_x_m_filter_hSS, data_x_m_filter_hSS_g, data_x_m_filter_hSS_n, data_x_m_filter_hSS_w)
```