---
title: \vspace{-1cm} Part3_B_Raw_DataProcessing_Filter_ECM \vspace{-0.20cm}
author: "Julien G. Roth"
date: "3/14/2021"
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
output: pdf_document
---

# Filtering for ECM Genes of Interest #

**The following script will output four dataframes which are subsets of the original processed dataframe. The 1st/2nd dataframe has been subset according to the 'Matrisome' list of ECM genes published by Nada and Hayes while the 3rd/4th dataframe has been subset by a list of neural ECM genes identified by JGR.**

The input are the dataframes: **data_x_m', 'data_x_m_f' (time-filtered).**

The output are the dataframes: **'data_x_m_ECM', 'data_x_m_f_ECM', 'data_x_m_neuralECM', 'data_x_m_f_neuralECM'.**

**Loading all ECM genes**

```{r, tidy=TRUE}

folder_ECM <- "ECMGeneIDs/"

mc_collagen <- read.table(str_c(directory, folder_ECM, "matrisome_core_collagen.txt"), quote="\"", comment.char="")
mc_collagen <- mutate(mc_collagen, rep("Collagen", nrow(mc_collagen)))
colnames(mc_collagen) <- c("Symbol", "Gene_Category")

mc_glycoproteins <- read.table(str_c(directory, folder_ECM, "matrisome_core_glycoproteins.txt"), quote="\"", comment.char="")
mc_glycoproteins <- mutate(mc_glycoproteins, rep("Glycoprotein", nrow(mc_glycoproteins)))
colnames(mc_glycoproteins) <- c("Symbol", "Gene_Category")

mc_proteoglycans_hyaluronan <- read.table(str_c(directory, folder_ECM, "matrisome_core_proteoglycans-hyaluronan.txt"),  quote="\"", comment.char="")
mc_proteoglycans_hyaluronan <- mutate(mc_proteoglycans_hyaluronan, rep("ProteoglycanHA", nrow(mc_proteoglycans_hyaluronan)))
colnames(mc_proteoglycans_hyaluronan) <- c("Symbol", "Gene_Category")

ma_affiliated <- read.table(str_c(directory, folder_ECM, "matrisome_associated_ecmaffiliatedproteins.txt"),  quote="\"", comment.char="")
ma_affiliated <- mutate(ma_affiliated, rep("ECMAffiliatedProtein", nrow(ma_affiliated)))
colnames(ma_affiliated) <- c("Symbol", "Gene_Category")

ma_regulators <- read.table(str_c(directory, folder_ECM, "matrisome_associated_regulators.txt"),  quote="\"", comment.char="")
ma_regulators <- mutate(ma_regulators, rep("Regulator", nrow(ma_regulators)))
colnames(ma_regulators) <- c("Symbol", "Gene_Category")

ma_integrins <- read.table(str_c(directory, folder_ECM, "matrisome_associated_integrins.txt"),  quote="\"", comment.char="")
ma_integrins <- mutate(ma_integrins, rep("Integrin", nrow(ma_integrins)))
colnames(ma_integrins) <- c("Symbol", "Gene_Category")

ma_secreted <- read.table(str_c(directory, folder_ECM, "matrisome_associated_secretedfactors.txt"),  quote="\"", comment.char="")
ma_secreted <- mutate(ma_secreted, rep("SecretedFactor", nrow(ma_secreted)))
colnames(ma_secreted) <- c("Symbol", "Gene_Category")

ECM <- rbind(mc_collagen, mc_glycoproteins, mc_proteoglycans_hyaluronan, ma_affiliated, ma_regulators, ma_integrins, ma_secreted)

nrow(ECM)
```

**Loading the neural ECM genes**

```{r, tidy=TRUE}

collagen <- read.table(str_c(directory, folder_ECM, "Collagen.txt"), quote="\"", comment.char="")
collagen <- mutate(collagen, rep("Collagen", nrow(collagen)))
colnames(collagen) <- c("Symbol", "Gene_Category")

enzyme <- read.table(str_c(directory, folder_ECM, "Enzyme.txt"), quote="\"", comment.char="")
enzyme <- mutate(enzyme, rep("Enzyme", nrow(enzyme)))
colnames(enzyme) <- c("Symbol", "Gene_Category")

fibronectin <- read.table(str_c(directory, folder_ECM, "Fibronectin.txt"), quote="\"", comment.char="")
fibronectin <- mutate(fibronectin, rep("Fibronectin", nrow(fibronectin)))
colnames(fibronectin) <- c("Symbol", "Gene_Category")

hyaluronicacid <- read.table(str_c(directory, folder_ECM, "HyaluronicAcid.txt"), quote="\"", comment.char="")
hyaluronicacid <- mutate(hyaluronicacid, rep("Hyaluronic Acid", nrow(hyaluronicacid)))
colnames(hyaluronicacid) <- c("Symbol", "Gene_Category")

integrin <- read.table(str_c(directory, folder_ECM, "Integrin.txt"),  quote="\"", comment.char="")
integrin <- mutate(integrin, rep("Integrin", nrow(integrin)))
colnames(integrin) <- c("Symbol", "Gene_Category")

laminin <- read.table(str_c(directory, folder_ECM, "Laminin.txt"),  quote="\"", comment.char="")
laminin <- mutate(laminin, rep("Laminin", nrow(laminin)))
colnames(laminin) <- c("Symbol", "Gene_Category")

linkprotein <- read.table(str_c(directory, folder_ECM, "LinkProteins.txt"),  quote="\"", comment.char="")
linkprotein <- mutate(linkprotein, rep("Link Protein", nrow(linkprotein)))
colnames(linkprotein) <- c("Symbol", "Gene_Category")

neuronalpentraxin <- read.table(str_c(directory, folder_ECM, "NeuronalPentraxin.txt"),  quote="\"", comment.char="")
neuronalpentraxin <- mutate(neuronalpentraxin, rep("Neuronal Pentraxin", nrow(neuronalpentraxin)))
colnames(neuronalpentraxin) <- c("Symbol", "Gene_Category")

proteoglycan <- read.table(str_c(directory, folder_ECM, "Proteoglycan.txt"),  quote="\"", comment.char="")
proteoglycan <- mutate(proteoglycan, rep("Proteoglycan", nrow(proteoglycan)))
colnames(proteoglycan) <- c("Symbol", "Gene_Category")

reelin <- read.table(str_c(directory, folder_ECM, "Reelin.txt"),  quote="\"", comment.char="")
reelin <- mutate(reelin, rep("Reelin", nrow(reelin)))
colnames(reelin) <- c("Symbol", "Gene_Category")

tenascin <- read.table(str_c(directory, folder_ECM, "Tenascin.txt"),  quote="\"", comment.char="")
tenascin <- mutate(tenascin, rep("Tenascin", nrow(tenascin)))
colnames(tenascin) <- c("Symbol", "Gene_Category")

thrombospondin <- read.table(str_c(directory, folder_ECM, "Thrombospondin.txt"),  quote="\"", comment.char="")
thrombospondin <- mutate(thrombospondin, rep("Thrombospondin", nrow(thrombospondin)))
colnames(thrombospondin) <- c("Symbol", "Gene_Category")

neuralECM <- rbind(collagen, enzyme, fibronectin, hyaluronicacid, integrin, laminin, linkprotein, neuronalpentraxin, proteoglycan, reelin, tenascin, thrombospondin)

nrow(neuralECM)
```


\pagebreak


**Filtering data based on all ECM genes**

```{r, tidy=TRUE}

data_x_m_ECM_merge <- data.frame(data_x_m[data_x_m$Symbol %in% ECM$Symbol, ])
data_x_m_ECM_categories <- data.frame(ECM[data_x_m_ECM_merge$Symbol %in% ECM$Symbol, ])

data_x_m_ECM_label <- merge.data.frame(data_x_m_ECM_merge, data_x_m_ECM_categories, by = "Symbol")
data_x_m_ECM_label$Full_Gene <- paste(data_x_m_ECM_label$Gene_Category, data_x_m_ECM_label$Isoform, sep = "-")

data_x_m_ECM <- data_x_m_ECM_label

data_x_m_ECM <- data_x_m_ECM[ , c(2, 3, 4, 1, 5, 6, 7, 8, 9)]

head(data_x_m_ECM, 5)
```

**Filtering time-subset data based on all ECM genes**

```{r, tidy=TRUE}

data_x_m_f_ECM_merge <- data.frame(data_x_m_f[data_x_m_f$Symbol %in% ECM$Symbol, ])
data_x_m_f_ECM_categories <- data.frame(ECM[data_x_m_f_ECM_merge$Symbol %in% ECM$Symbol, ])

data_x_m_f_ECM_label <- merge.data.frame(data_x_m_f_ECM_merge, data_x_m_f_ECM_categories, by = "Symbol")
data_x_m_f_ECM_label$Full_Gene <- paste(data_x_m_f_ECM_label$Gene_Category, data_x_m_f_ECM_label$Isoform, sep = "-")

data_x_m_f_ECM <- data_x_m_f_ECM_label

data_x_m_f_ECM <- data_x_m_f_ECM[ , c(2, 3, 4, 8, 9, 1, 5, 6, 7, 10, 11)]

head(data_x_m_f_ECM, 5)
```

**Filtering data based on neural ECM genes**

```{r, tidy=TRUE}

data_x_m_neuralECM_merge <- data.frame(data_x_m[data_x_m$Symbol %in% neuralECM$Symbol, ])
data_x_m_neuralECM_categories <- data.frame(neuralECM[data_x_m_neuralECM_merge$Symbol %in% neuralECM$Symbol, ])

data_x_m_neuralECM_label <- merge.data.frame(data_x_m_neuralECM_merge, data_x_m_neuralECM_categories, by = "Symbol")
data_x_m_neuralECM_label$Full_Gene <- paste(data_x_m_neuralECM_label$Gene_Category, data_x_m_neuralECM_label$Isoform, sep = "-")

data_x_m_neuralECM <- data_x_m_neuralECM_label

data_x_m_neuralECM <- data_x_m_neuralECM[ , c(2, 3, 4, 1, 5, 6, 7, 8, 9)]

head(data_x_m_neuralECM, 5)
```

**Filtering time-subset data based on neural ECM genes**

```{r, tidy=TRUE}

data_x_m_f_neuralECM_merge <- data.frame(data_x_m_f[data_x_m_f$Symbol %in% neuralECM$Symbol, ])
data_x_m_f_neuralECM_categories <- data.frame(neuralECM[data_x_m_f_neuralECM_merge$Symbol %in% neuralECM$Symbol, ])

data_x_m_f_neuralECM_label <- merge.data.frame(data_x_m_f_neuralECM_merge, data_x_m_f_neuralECM_categories, by = "Symbol")
data_x_m_f_neuralECM_label$Full_Gene <- paste(data_x_m_f_neuralECM_label$Gene_Category, data_x_m_f_neuralECM_label$Isoform, sep = "-")

data_x_m_f_neuralECM <- data_x_m_f_neuralECM_label

data_x_m_f_neuralECM <- data_x_m_f_neuralECM[ , c(2, 3, 4, 8, 9, 1, 5, 6, 7, 10, 11)]

head(data_x_m_f_neuralECM, 5)
```


\pagebreak


# Cleaning Data #

**Export the variables which are requried for downstream steps.**

```{r, tidy=TRUE}

write.table(data_x_m_ECM, str_c(directory, variables, "data_x_m_ECM.csv"), sep = ',', row.names = FALSE)
write.table(data_x_m_f_ECM, str_c(directory, variables, "data_x_m_f_ECM.csv"), sep = ',', row.names = FALSE)
write.table(data_x_m_neuralECM, str_c(directory, variables, "data_x_m_neuralECM.csv"), sep = ',', row.names = FALSE)
write.table(data_x_m_f_neuralECM, str_c(directory, variables, "data_x_m_f_neuralECM.csv"), sep = ',', row.names = FALSE)
```

**Remove less relevant or temporary variables to decrease clutter.**

```{r, tidy=TRUE}

rm(collagen, data_x_m_ECM_categories, data_x_m_ECM_label, data_x_m_ECM_merge, data_x_m_f_ECM_categories, data_x_m_f_ECM_label, data_x_m_f_ECM_merge, data_x_m_f_neuralECM_categories, data_x_m_f_neuralECM_label, data_x_m_f_neuralECM_merge, data_x_m_neuralECM_categories, data_x_m_neuralECM_merge, data_x_m_neuralECM_label, enzyme, fibronectin, hyaluronicacid, integrin, laminin, linkprotein, ma_affiliated, ma_regulators, ma_secreted, mc_collagen, mc_proteoglycans_hyaluronan, mc_glycoproteins, ma_integrins, neuronalpentraxin, proteoglycan, reelin, tenascin, thrombospondin)
```