---
title: \vspace{-1cm} Part6_Raw_DataProcessing_DESEQformat \vspace{-0.20cm}
author: "Julien G. Roth"
date: "3/14/2021"
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
output: pdf_document
---

# Formatting Raw Data to Run in DESEQ2 #

The data is rearragned to create a read counts table and a metadata table.

Input: **data_x_m_f_neuralECM**

Output: **data_x_m_f_neuralECM_log_f**

Key: x (Trevino study), m (mean expression), f (timepoint filtered), neuralECM (neural ECM gene filtered), 

```{r, tidy=TRUE, message=FALSE, warning=FALSE}

data_x_m_f_neuralECM_log <- data_x_m_f_neuralECM

data_x_m_f_neuralECM_log[data_x_m_f_neuralECM_log == 0] <- -999

data_x_m_f_neuralECM_log <- data_x_m_f_neuralECM_log %>% 
  mutate(Transformed = log(Mean_Expression, base = 2))

data_x_m_f_neuralECM_log_f <- data_x_m_f_neuralECM_log[!is.na(data_x_m_f_neuralECM_log$Transformed),]
```

```{r, tidy=TRUE}

nrow(data_x_m_f_neuralECM_log_f)
```

```{r, tidy=TRUE}

head(data_x_m_f_neuralECM_log_f, 5)
```


\pagebreak


# Cleaning Data #

**Export the variables which are requried for downstream steps.**

```{r, tidy=TRUE}

write.table(data_x_m_f_neuralECM_log_f, str_c(directory, variables, "data_x_m_f_neuralECM_log_f.csv"), sep = ',', row.names = FALSE)
```

**Remove less relevant or temporary variables to decrease clutter.**

```{r, tidy=TRUE}

rm(data_x_m_f_neuralECM_log)
```