---
title: \vspace{-1cm} Part6_Raw_DataProcessing_DESEQformat \vspace{-0.20cm}
author: "Julien G. Roth"
date: "3/14/2021"
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
output: pdf_document
---

# Formatting Raw Data to Run in DESEQ2 #

The data is rearragned to create a read counts table and a metadata table. It seems that using the log_f data frame created in Part 5 results in multiple genes/samples having NAs. While helpful, it is not requried to filter low expression genes for DESEQ. Given the low expression (and small number) of neuralECM genes, the following script (and subsequent DESEQ analysis) does not use expression filted data.

Input: **data_x_m_f_neuralECM_log_f** OR **data_x_m_f_neuralECM**

Output: **data_x_m_f_neuralECM_f_counts**, **data_x_m_f_neuralECM_metadata**

Key: x (Trevino study), m (mean expression), f (timepoint filtered), neuralECM (neural ECM gene filtered), f (low expression filtered)

**Creating the counts table**

```{r, tidy=TRUE}

counts <- data_x_m_f_neuralECM[, c(1:4, 11, 8)]

counts <- counts %>% 
  mutate(Comp = paste(Patterning, Cell_Type, Day, Maturity, sep = "-"))

counts <- counts[, c(1:4, 7, 5:6)]

counts_table <- counts[, c(6,5,7)]

counts_table_spread <- counts_table %>% 
  spread(Comp, Mean_Expression) 

rownames(counts_table_spread) <- counts_table_spread$Full_Gene
counts_table_spread <- counts_table_spread[, 2:20]

data_x_m_f_neuralECM_f_counts <- counts_table_spread

colnames(data_x_m_f_neuralECM_f_counts)
```

**Creating the metadata table**

```{r, tidy=TRUE}

meta <- counts[, c(5, 1:4)]

meta_f <- meta[!duplicated(meta), ]

meta_f <- meta_f[order(meta_f$Comp), ]

data_x_m_f_neuralECM_metadata <- meta_f

data_x_m_f_neuralECM_metadata$Comp
```

```{r, tidy=TRUE}
all(data_x_m_f_neuralECM_metadata$Comp == colnames(data_x_m_f_neuralECM_f_counts))
```


\pagebreak


# Cleaning Data #

**Export the variables which are requried for downstream steps.**

```{r, tidy=TRUE}

write.table(data_x_m_f_neuralECM_f_counts, str_c(directory, variables, "data_x_m_f_neuralECM_f_counts.csv"), sep = ',', row.names = TRUE)
write.table(data_x_m_f_neuralECM_metadata, str_c(directory, variables, "data_x_m_f_neuralECM_metadata.csv"), sep = ',', row.names = FALSE)
```

**Remove less relevant or temporary variables to decrease clutter.**

```{r, tidy=TRUE}

rm(counts, counts_table, counts_table_spread, meta, meta_f)
```